### 2020-07-20 paper/paper.tex
Initial commit.

% !TEX TS-program = xelatex
%
% Created by Chris on 2020-07-19.
% Copyright (c) Chris von Csefalvay, 2020.
\documentclass{article}
\usepackage{amsmath}
\usepackage{polyglossia}
\usepackage{hyperref}

% Bibliography styling
\usepackage[super,square,sort&compress,numbers]{natbib}
\bibliographystyle{unsrtnat}
\usepackage{url}
\urlstyle{same}

\hypersetup
{
  pdftitle   = {A differential game analysis of social distancing in COVID-19},
  pdfauthor  = {Chris von Csefalvay}
}

\title{A differential game analysis of social distancing in COVID-19}
\author{Chris von Csefalvay}

\begin{document}

\maketitle

\begin{abstract}
    Abstract
\end{abstract}

\section{Introduction} % (fold)
\label{sec:introduction}
Where an infectious disease is not amenable to population-level prevention through vaccination and risks are non-trivial, non-pharmaceutical interventions (NPIs) remain the principal tool of public health to respond to an outbreak. This is the case with novel infectious diseases that have no specific treatment and no prophylactic (vaccine) available. In the absence of pharmaceutical interventions of proven effectiveness, in particular prophylactically, the main public health response to the emerging pandemic of COVID-19, a viral syndrome caused by the (+)ssRNA virus SARS-CoV-2 (order \emph{Nidovirales}, family \emph{Coronaviridae}, genus \emph{Betacoronavirus}, subgenus \emph{Sarbecovirus}), has rested principally on NPIs, first and foremost social distancing and ancillary steps intended to facilitate that.

In general, any course of conduct that reduces the encounter rate between an individual and other individuals can be considered a form of social distancing. This may be brought about through limiting public facilities for such encounters ('lockdowns'), through limiting individual gatherings by size ('large-gathering bans') and through encouraging individual social distancing. From the perspective of game theory, social distancing can be viewed as a non-cooperative game of a population $P_{1 \ldot n}$ of size $n$, where at any given time $t \in [t_0, t_{e}]$, the strategy adopted by $p_i$ is denoted as $\sigma (p_i, t)$. For the sake of simplicity, we assume that distancing is either not exercised at all or perfectly exercised, i.e. $\sigma (p_i, t) \in \{ 0, 1 \}$. Then, for the entire population, the overall strategy can be described as

\begin{equation}
	\bar{\sigma}(P, t) = \frac{\displaystyle \sum_{i = 0}^n \sigma(p_i, t)}{n}
	\label{eq:overall_strategy}
\end{equation}

\noindent and for the entire time period, from $t_0$ to the endpoint $t_e$ (which may be eradication, elimination, natural extinction of the pathogen, the availability of a vaccine or a combination thereof), for discrete time $t$,

\begin{equation}
	\bar{\bar{\sigma}}(P) = \frac{\displaystyle \sum_{i = 0}^n \displaystyle \sum_{j = 0}^{t_e - t_0} \sigma(p_i, j)}{n (t_e - t_0)}
	\label{eq:overall_strategy_over_time}
\end{equation}

However, with each course of conduct, there is associated a cost $J(\sigma)$. For simplicity's sake, let us consider these costs to be governed by the following three precepts.

\begin{enumerate}
	\item A person $p_i \in P$ opting for strategy $1$ (social distancing) will incur $c_d$, the immediate costs of distancing. These may be social (lessened social interaction), psychological (lessened access to support systems), economic (lower access to facilities to earn) or simple matters of convenience (access to amenities). While $c_d$ is somewhat dependent on $\bar{\bar{(P)}}$ (thus not distancing does not yield a benefit to a lone social distancer in terms of access to amenities if all of the latter are closed), it can be assumed to be largely constant.
	\item A person $p_i \in P$ opting for strategy $1$ will gain a benefit of $b_d$, which represents an individual's reduced risk of illness due to their own distancing from possibly infected persons.
	\item Regardless of their individual strategy, a person $p_i \in P$ will gain a benefit $b_s$ that is very closely approximated by a function of the overall strategy adopted by society, so that $b_s(p_i, t) = f(\bar{\sigma}(P, t))$ for $p_i \in P$.
\end{enumerate}

Then, assigning the variable $c_s$ to represent the cost of illness (economic loss, medical costs, long-term health risks), we can express the cost functions of each strategy, $\sigma_1$ (social distancing) and $\sigma_0$ (no social distancing), as

\begin{equation}
	J_{\sigma_1}(p_i, t) = c_d - b_d r_d c_s - f(\bar{\sigma}(\delta_P, t))
\end{equation}

\noindent and

\begin{equation}
	J_{\sigma_0}(p_i, t) = r_d c_s - f(\bar{\sigma}(\delta_P, t))
\end{equation}

\noindent for $p_i \in P$, where $r_d$ denotes $p$'s level of risk and $\delta_P$ denotes the proportion of individuals that adopt $\sigma_1$. And consequently, it holds for the. population level cost $\bar{J}$ that

\begin{equation}
	\bar{J}(P, t) = \frac{\delta_P J_{\sigma_1}(P, t) + (1 - \delta_P) J_{\sigma_0}(P, t)}{n}
\end{equation}

As Reluga (2010) notes, the effect of social distancing diminishes with the increase in participants, the phenomenon of diminishing returns.\cite{reluga2010game} Thus, as the number of individuals in $P$ opting for social distancing increases, the marginal increase by another social distancer diminishes. This can be represented by a discount factor $h$, to yield

\begin{equation}
	\bar{J}(P, t) = e^{-ht} \Big( \frac{\delta_P J_{\sigma_1}(P, t) + (1 - \delta_P) J_{\sigma_0}(P, t)}{n} \Big)
\end{equation}

However, because individual decisions affect the overall gain (due to the component dependent on $\bar{\sigma}(P, t)$), our principal concern is not with individual action but with analysis of such strategies on a population level. This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution} and, since its publication in 1927, widely adapted and adopted.\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization}

% section introduction (end)

\section{Methods} % (fold)
\label{sec:methods}

Given a population of $n$ under the assumption that reinfection is impossible (as, e.g., in the case of measles) or rare (as is the case for SARS-CoV-2\cite{edridge2020human,deng2020primary,bao2020reinfection}), and neglecting for the time being the vital dynamics (birth, unrelated death, migration) of the population, the dynamics of any population can be modelled as a system of ordinary differential equations

\begin{equation}
	\begin{aligned}
		\frac{dS}{dt} = - \frac{\beta S I}{n} 								\\
		\frac{dI}{dt} = \frac{\beta S I}{n} - \gamma I 						\\
		\frac{dR}{dt} = \gamma I
	\end{aligned}
	\label{eq:sir_equation}
\end{equation}

\noindent under the assumption of $S + I + R = 0$, where $S$ represents susceptible individuals, $I$ represents infected/infectious individuals and $R$ accounts for removed individuals (mortality and recovery to immunity). Even in the absence of firm evidence as to whether SARS-CoV-2 infection followed by recovery would engender lifelong immunity or not,\cite{roy2020covid,ota2020will,lin2020duration} it can be assumed in the short term -- based on evidence from MERS-CoV and SARS-CoV -- that in the short term, survivors remain immune,\cite{prompetchara2020immune} and consequently the number of individuals in $R$ does not decrease.

For a population $P$ of size $n$ that pursues the population-level aggregate of strategies described by $\bar{\sigma(P)}$, the dynamics described above in Equation~\eqref{eq:sir_equation} becomes

\begin{equation}
	\begin{aligned}
		\frac{dS}{dt} = - \bar{\sigma}(P) \frac{\beta S I}{n} 				\\
		\frac{dI}{dt} = \bar{\sigma}(P) \frac{\beta S I}{n} - \gamma I 		\\
		\frac{dR}{dt} = \gamma I
	\end{aligned}
	\label{eq:sir_strat_equation}
\end{equation}

The fraction $\frac{\beta n}{\gamma}$ equals the basic reproduction number, $R_0$. For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive}

% section methods (end)

\bibliography{bibliography}

\end{document}

### 2020-07-21 paper/paper.tex
Notebook on ODE added.

@@
 \urlstyle{same}
+
+% Language and hyphenation
+% \usepackage[english]{babel}
+\usepackage[htt]{hyphenat}
+
+% Graphics
+\usepackage{graphicx}
 
@@
 
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig1-odes}
+	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000.
+For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
+(2020),\cite{liu2020reproductive} and the $\gamma$ parameter.
+The susceptible population is displayed in blue, while infected/infectious cases are marked in red and recovered cases in green.}
+	\label{fig:ode_solutions}
+\end{figure}
+
 The fraction $\frac{\beta n}{\gamma}$ equals the basic reproduction number, $R_0$.
-For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive}
+For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the invers of the number of days of illness, which can be approximated as $8.5 \pm 3.9$ days.
+Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
+
+The ODEs of Equation~\eqref{eq:sir_equation} can then be parametrised with the population-level strategy (Equation~\eqref{eq:sir_strat_equation}).
 

### 2020-07-21 paper/paper.tex
Added parts on cost-risk equilibria.

--- 2020-07-20 paper/paper.tex
@@
 
-However, with each course of conduct, there is associated a cost $J(\sigma)$. For simplicity's sake, let us consider these costs to be governed by the following three precepts.
+However, with each course of conduct, there is associated a cost $J(\sigma)$.
+For simplicity's sake, let us consider these costs to be governed by the following three precepts given a pathogen with fixed characteristics ($R_0$, transmission potential, risks \&c.).
+Let $\bar{\delta}(P)$ equal the proportion of persons $p \in P$ adopting strategy $\sigma_{\delta}$.
+It then holds that:
 
 \begin{enumerate}
-	\item A person $p_i \in P$ opting for strategy $1$ (social distancing) will incur $c_d$, the immediate costs of distancing.
+	\item A person $p_i \in P$ opting for strategy $\delta$ (social distancing) will incur $c_d$, the immediate costs of distancing.
 These may be social (lessened social interaction), psychological (lessened access to support systems), economic (lower access to facilities to earn) or simple matters of convenience (access to amenities).
 While $c_d$ is somewhat dependent on $\bar{\bar{(P)}}$ (thus not distancing does not yield a benefit to a lone social distancer in terms of access to amenities if all of the latter are closed), it can be assumed to be largely constant.
-	\item A person $p_i \in P$ opting for strategy $1$ will gain a benefit of $b_d$, which represents an individual's reduced risk of illness due to their own distancing from possibly infected persons.
-	\item Regardless of their individual strategy, a person $p_i \in P$ will gain a benefit $b_s$ that is very closely approximated by a function of the overall strategy adopted by society, so that $b_s(p_i, t) = f(\bar{\sigma}(P, t))$ for $p_i \in P$.
+	\item Compared to a person opting for strategy $\delta$, a person $p_i \in P$ opting for strategy $\lnot \delta$ will incur a relative additional risk $f_r(\bar{\delta}(P), t)$, i.e.
+contingent on the population's behaviour at time $t$.
+	\item Finally, a person $p_i \in P$ will regardless of their individual choice (at least at non-trivial population sizes) gain a protective benefit, which is a function of the population's adherence to social distancing.
+We will denote this term as $f_c(\bar{\delta}(P), t)$.
 \end{enumerate}
 
-Then, assigning the variable $c_s$ to represent the cost of illness (economic loss, medical costs, long-term health risks), we can express the cost functions of each strategy, $\sigma_1$ (social distancing) and $\sigma_0$ (no social distancing), as
+Then, assigning the variable $c_s$ to represent the cost of illness (economic loss, medical costs, long-term health risks), we can express the cost functions of each strategy, $\sigma_{\delta}$ (social distancing) and $\sigma_{\lnot \delta}$ (no social distancing), as
 
 \begin{equation}
-	J_{\sigma_1}(p_i, t) = c_d - b_d r_d c_s - f(\bar{\sigma}(\delta_P, t))
+	J_{\sigma_{\delta}}(p_i, t) = c_d - f_c(\bar{\delta}(P), t)
 \end{equation}
@@
 \begin{equation}
-	J_{\sigma_0}(p_i, t) = r_d c_s - f(\bar{\sigma}(\delta_P, t))
+	J_{\sigma_{\lnot \delta}}(p_i, t) = f_r(\bar{\delta}(P), t) c_s - f_c(\bar{\delta}(P), t)
 \end{equation}
 
-\noindent for $p_i \in P$, where $r_d$ denotes $p$'s level of risk and $\delta_P$ denotes the proportion of individuals that adopt $\sigma_1$.
-And consequently, it holds for the.
-population level cost $\bar{J}$ that
+\noindent And consequently, it holds for the population level cost $\bar{J}$ at time $t$ that
 
 \begin{equation}
-	\bar{J}(P, t) = \frac{\delta_P J_{\sigma_1}(P, t) + (1 - \delta_P) J_{\sigma_0}(P, t)}{n}
+	\bar{J}(P, t) = \frac{\displaystyle \sum_{i=0}^{\delta(P) n} J_{\sigma_{\delta}}(p_i, t) + \displaystyle \sum_{j=0}^{1-\delta(P) n} J_{\sigma_{\lnot \delta}}(p_j, t)}{n}
+	\label{eq:cost_eqn}
+\end{equation}
+
+\noindent and for a population over time $t_0$ to $t_e$,
+
+\begin{equation}
+	\bar{\bar{J}}(P) = \int_{t_0}^{t_e} \bar{J}(P, t) dt
 \end{equation}
@@
 \begin{equation}
-	\bar{J}(P, t) = e^{-ht} \Big( \frac{\delta_P J_{\sigma_1}(P, t) + (1 - \delta_P) J_{\sigma_0}(P, t)}{n} \Big)
+	\bar{J}(P) = \int_{t_0}^{t_e} e^{-ht} \frac{\delta(P) J_{\sigma_{\delta}}(P, t) + (1 - \delta(P)) J_{\sigma_{\lnot \delta}}(P, t)}{n}
 \end{equation}
 
-However, because individual decisions affect the overall gain (due to the component dependent on $\bar{\sigma}(P, t)$), our principal concern is not with individual action but with analysis of such strategies on a population level.
+However, because individual decisions affect the overall gain (due to the component dependent on $\delta(P, t)$), our principal concern is not with individual action but with analysis of such strategies on a population level.
 This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution} and, since its publication in 1927, widely adapted and adopted.\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization}
@@
 \label{sec:methods}
+
+\subsection{The ordinary differential equations of disease dynamics} % (fold)
+\label{sub:the_ordinary_differential_equations_of_disease_dynamics}
 
@@
 
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig1-odes}
+	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000 and a seed population of 1\% infected initially.
+For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
+(2020),\cite{liu2020reproductive} and the $\gamma$ parameter.
+The susceptible population is displayed in blue, while infected/infectious cases are marked in red and recovered cases in green.}
+	\label{fig:ode_solutions}
+\end{figure}
+
 The fraction $\frac{\beta n}{\gamma}$ equals the basic reproduction number, $R_0$.
-For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive}
+For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the invers of the number of days of illness, which can be approximated as $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
+
+% subsection the_ordinary_differential_equations_of_disease_dynamics (end)
+
+\subsection{Population strategy contingent solutions to population dynamics} % (fold)
+\label{sub:population_strategy_contingent_solutions_to_population_dynamics}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig2-strategy_solutions_by_days}
+	\caption{A first naive approximation of population strategy contingent dynamics in a population of 10,000 with an initial seed population of 1\% infected initially, showing the maximum number of infected by 14, 21, 28 and 35 days.
+For each plot, the $\gamma$ parameter was estimated to be $\frac{1}{8.5}$, based on Pan et al.
+(2020) and Liu et al.
+(2020),\cite{pan2020clinical,liu2020risk}.
+This initial approach assumes a constant effect to social distancing, i.e.
+not accounting for diminishing returns, whether over time or over population participating in social distancing.}
+	\label{fig:first_naive_solution}
+\end{figure}
+
+Let us now draw on Equation~\eqref{eq:overall_strategy} to adjust the system of differential equations in Equation~\eqref{eq:sir_equation}.
+A population $P$ adopting the overall strategy $\bar{\sigma}(P, t)$ at time $t$ will have an efficiency factor of $\delta(P, t)$, bounded by $[0; 1]$, where
+
+\begin{equation}
+	\begin{aligned}
+		\delta(P, t) = \frac{\displaystyle \sum_{i = 1}^n \sigma(p_i, t)}{n}
+	\end{aligned}
+	\label{eq:delta_p_t}
+\end{equation}
+
+\noindent in other words, $\delta(P, t)$ is the fraction of the population $P$ that is engaged in social distancing at time $t$.
+An initial naive approximation of the effect of various values of $\delta(P, t)$ (as a rough proxy of $\bar{\sigma{(P, t)}}$) is represented by Figure~\ref{fig:first_naive_solution}.
+This does not, however, account for the underlying differential game, i.e.
+the individual and societal cost of distancing or not distancing.
+In the following refinement of the model, we will consider each of these in turn.
+
+% subsection population_strategy_contingent_solutions_to_population_dynamics (end)
+
+\subsection{Cost and strategy} % (fold)
+\label{sub:cost_and_strategy}
+
+For each member $p$ of the population $P$, there exists a cost function of distancing $J_{\delta}(P, t)$ for any given time $t$ contingent on the population $P$'s behaviour, and a corresponding function $J_{\lnot \delta}(P, t)$ for not distancing (see Equation~\eqref{eq:cost_eqn}).
+As noted above, three major factors govern the cost for each individual: the cost of distancing per unit time ($c_d$, comprising economic and non-economic costs of social distancing), the risk of contracting the disease ($f_r(\bar{\delta}(P), t)$) and the cost of sickness per unit time spent sick ($c_s$).
+The cost is decreased by the social benefit of distancing, which is, at least outside trivially sized populations, largely independent from the individual and depends on how many in the population are engaging in social distancing ($f_c(\bar{\delta}(P), t)$).
+
+For a population of rational actors, it can be assumed that they will adopt the strategy that is the most beneficial given the information they have.
+It deserves mention that, as Reluga (2006) alludes to, decisions are made by players based on the information they have,\cite{reluga2006evolving} which may often be incomplete or even outright false: perceptions, not necessarily reality, drive decision-making.
+Ignoring the social benefit, the equilibrium point between strategies, $J_{\delta}(P, t) = J_{\lnot \delta}(P, t)$ is reached for the equilibrium condition where
+
+\begin{equation}
+	c_d = f_r(\bar{\delta}(P), t) c_s
+	\label{eq:cost_risk}
+\end{equation}
+
+\noindent -- or in other words, where the cost of distancing equals the risk of contracting the disease (given the population's state) and the cost of the disease.
+This point also represents a subgame perfect Nash equilibrium for the underlying differential game.
+This is easy to see: given this equilibrium point, a person who is not socially distancing may switch to the strategy of social distancing, but would incur an unwarranted cost.
+Similarly, a person who is socially distancing at this time would risk more than the costs avoided by forgoing social distancing.
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig3-risk-of-adherence}
+	\caption{Relative benefits of social distancing as a function of $R_0$ and societal adherence for a population of 10,000 with an initial infected seed population of 1\%.
+Benefits are normalised to 1.0 against the maximum value.
+For each plot, the $\gamma$ parameter was estimated to be $\frac{1}{8.5}$, based on Pan et al.
+(2020) and Liu et al.
+(2020),\cite{pan2020clinical,liu2020risk}.
+As the plot indicates, social distancing is most effective for higher $R_0$ and mediocre overall social distancing within the population, contradicting the widely shared popular perception that social distancing is useless unless practised by an overwhelming majority and supporting the case for social distancing for all individuals.}
+	\label{fig:risk_and_compliance}
+\end{figure}
+
+It is clear that the risk of disease at time $t$, $f_r(\bar{\delta}(P), t)$, is in fact reflective of the adjusted force of infection subject to the share of the population engaged in social distancing, described by $\bar{\delta}(P)$.
+The force of infection $\lambda$ is typically defined as $\frac{\beta I}{n}$.
+When adjusting for the risk of disease to an individual considering the population's response, encapsulated in $\bar{\delta}(P)$, we get
+
+\begin{equation}
+	f_r(\bar{\delta}(P), t) = \frac{\bar{\delta}(P) \beta I(t)}{n}
+	\label{eq:adjusted_force_of_infection}
+\end{equation}
+
+\noindent and substituting $R_0 = \frac{\beta n}{\gamma}$ into Equation~\eqref{eq:adjusted_force_of_infection} allows us to express this as
+
+\begin{equation}
+	f_r(\bar{\delta}(P), t) = \frac{\bar{\delta}(P) R_0 \gamma I(t)}{n^2}
+	\label{eq:expanded_cost_fx}
+\end{equation}
+
+This can be solved using quantitative methods (see Figure~\ref{fig:risk_and_compliance}) to yield isorisk curves based on $R_0$ and $\delta(P)$, i.e.
+on pathogen-intrinsic and population-intrinsic characteristics, over time.
+
+% subsection cost_and_strategy (end)
+
+\subsection{Cost-risk equilibria} % (fold)
+\label{sub:cost_risk_equilibria}
+
+As noted in Equation~\eqref{eq:cost_risk}, on a social level, the equilibrium state is defined where the cost of disease times the risk of disease equals the cost of avoidance.
+By substituting Equation~\eqref{eq:expanded_cost_fx} into Equation~\eqref{eq:cost_risk}, we get the equilibrium equation
+
+\begin{equation}
+	c_d(t) = c_s(t) \frac{\bar{\delta}(P(t)) R_0 \gamma I(t)}{n^2}
+\end{equation}
+
+\noindent or, reformulated,
+
+\begin{equation}
+	\frac{c_d(t)}{c_s(t)} = \frac{\bar{\delta}(P(t)) R_0 \gamma I(t)}{n^2}
+\end{equation}
+
+Expressed over the entire duration $t_{0 \ldots e}$, the cost ratio of distancing ($c_d$) and sickness ($c_s$) is
+
+\begin{equation}
+	\frac{c_d}{c_t} = \int_{t_0}^{t_e} \frac{\bar{\delta}(P(t)) R_0 \gamma I(t)}{n^2} dt
+\end{equation}
+
+% subsection cost_risk_equilibria (end)
 
 % section methods (end)
+
+\section{Results} % (fold)
+\label{sec:results}
+
+This risk, viewed as a function of $R_0$ and $\delta(P)$ in Figure~\ref{fig:risk_and_compliance}, indicates the somewhat counterintuitive finding that social distancing becomes more,
+
+% section results (end)
 

### 2020-07-23 paper/paper.tex
Result writeup.

--- 2020-07-20 paper/paper.tex
@@
 
+% Language and hyphenation
+% \usepackage[english]{babel}
+\usepackage[htt]{hyphenat}
+
+% Graphics
+\usepackage{graphicx}
+
 \hypersetup
 {
-  pdftitle   = {A differential game analysis of social distancing in COVID-19},
+  pdftitle   = {Marginal dynamics of social distancing in COVID-19},
   pdfauthor  = {Chris von Csefalvay}
@@
 
-\title{A differential game analysis of social distancing in COVID-19}
+\title{Marginal dynamics of social distancing in COVID-19}
 \author{Chris von Csefalvay}
@@
 \begin{equation}
-	J_{\sigma_0}(p_i, t) = r_d c_s - f(\bar{\sigma}(\delta_P, t))
+	J_{\sigma_{\lnot \delta}}(p_i, t) = f_r(\bar{\delta}(P), t) c_s - f_c(\bar{\delta}(P), t)
+	\label{eq:risk}
 \end{equation}
 
-\noindent for $p_i \in P$, where $r_d$ denotes $p$'s level of risk and $\delta_P$ denotes the proportion of individuals that adopt $\sigma_1$.
-And consequently, it holds for the.
-population level cost $\bar{J}$ that
+\noindent And consequently, it holds for the population level cost $\bar{J}$ at time $t$ that
 
 \begin{equation}
-	\bar{J}(P, t) = \frac{\delta_P J_{\sigma_1}(P, t) + (1 - \delta_P) J_{\sigma_0}(P, t)}{n}
+	\bar{J}(P, t) = \frac{\displaystyle \sum_{i=0}^{\delta(P) n} J_{\sigma_{\delta}}(p_i, t) + \displaystyle \sum_{j=0}^{1-\delta(P) n} J_{\sigma_{\lnot \delta}}(p_j, t)}{n}
+	\label{eq:cost_eqn}
+\end{equation}
+
+\noindent and for a population over time $t_0$ to $t_e$,
+
+\begin{equation}
+	\bar{\bar{J}}(P) = \int_{t_0}^{t_e} \bar{J}(P, t) dt
 \end{equation}
@@
 
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig1-odes}
+	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000 and a seed population of 1\% infected initially.
+For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
+(2020),\cite{liu2020reproductive} and the $\gamma$ parameter.
+The susceptible population is displayed in blue, while infected/infectious cases are marked in red and recovered cases in green.}
+	\label{fig:ode_solutions}
+\end{figure}
+
 \noindent under the assumption of $S + I + R = 0$, where $S$ represents susceptible individuals, $I$ represents infected/infectious individuals and $R$ accounts for removed individuals (mortality and recovery to immunity).
+The fraction $\frac{\beta}{\gamma}$ equals the basic reproduction number, $R_0$.
+For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the invers of the number of days of illness, which can be approximated as $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
 Even in the absence of firm evidence as to whether SARS-CoV-2 infection followed by recovery would engender lifelong immunity or not,\cite{roy2020covid,ota2020will,lin2020duration} it can be assumed in the short term -- based on evidence from MERS-CoV and SARS-CoV -- that in the short term, survivors remain immune,\cite{prompetchara2020immune} and consequently the number of individuals in $R$ does not decrease.
+Numerical solutions to this system of differential equations have been calculated using \texttt{odepack} via \texttt{SciPy 1.5.1}\cite{virtanen2020scipy} on Python 3.6, and are presented as Figure~\ref{fig:ode_solutions} for a range of values of $R_0$ and $\gamma$.
 
-For a population $P$ of size $n$ that pursues the population-level aggregate of strategies described by $\bar{\sigma(P)}$, the dynamics described above in Equation~\eqref{eq:sir_equation} becomes
+% subsection the_ordinary_differential_equations_of_disease_dynamics (end)
+
+\subsection{Population strategy contingent solutions to population dynamics} % (fold)
+\label{sub:population_strategy_contingent_solutions_to_population_dynamics}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/fig3-SIR-by-delta}
+	\caption{Contour diagrams of numeric solutions for susceptible, infectious and removed ($S$, $I$ and $R$) compartment sizes for a base population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
+As this figure indicates, the 'critical mass' of social distancing takes place in the $\delta$ range of 0 to 0.4, and thus even modest increases in social distancing participation at low levels can make a significant difference in the number of infectious cases.}
+	\label{fig:fig3-SIR-by-delta}
+\end{figure}
+
+Given a population that then adopts a sum strategy $\bar{\sigma}(P, t)$ at time $t$ that results in $\delta(P, t)$ adherence (or in Reluga's terms, investment) to social distancing, the flow from $S$ to $I$ is reduced by a corresponding factor.
+This allows us to rewrite Equation~\eqref{eq:sir_equation} so that for a society level strategy $\bar{\sigma}$ yielding $\delta$, the populations can be characterised as
 
@@
 	\begin{aligned}
-		\frac{dS}{dt} = - \bar{\sigma}(P) \frac{\beta S I}{n} 				\\
-		\frac{dI}{dt} = \bar{\sigma}(P) \frac{\beta S I}{n} - \gamma I 		\\
+		\frac{dS}{dt} = - \frac{\beta S I - \delta \beta S I}{n} 			\\
+		\frac{dI}{dt} = \frac{\beta S I - \delta \beta S I}{n} - \gamma I	\\
 		\frac{dR}{dt} = \gamma I
 	\end{aligned}
-	\label{eq:sir_strat_equation}
+	\label{eq:sir_with_social_distancing}
 \end{equation}
 
-The fraction $\frac{\beta n}{\gamma}$ equals the basic reproduction number, $R_0$.
-For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive}
+Solutions to this system of differential equations have been calculated and are presented in Figure~\ref{fig:fig3-SIR-by-delta}.
+Importantly, this allows us to identify the marginal utility $\hat{U}(P, t)$ as
+
+\begin{equation}
+	\begin{aligned}
+		\hat{U}(P, t) = \frac{\partial I(P, t)}{\partial \delta(t)}
+	\end{aligned}
+	\label{eq:marginal_utility}
+\end{equation}
+
+\noindent i.e.
+the partial derivative of $I(P, t)$ over $\delta(P, t)$.
+Thus, for a population-level strategy $\bar{\sigma}(P, t)$ associated with a compliance rate (i.e.
+the fraction of persons $p_i \in P$ engaged in social distancing at time $t$) of $\delta(P, t)$, there exists a function $\hat{U}(P, t)$ given $\gamma$ and $R_0$ that indicates the marginal utility at any given value of $\delta(P, t)$.
+This, too, can be numerically ascertained, and is shown on Figure~\ref{fig:marginal_utility}.
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/marginal_utility}
+	\caption{Marginal utility of social distancing in a population $P$ of $\delta(P)$ adherence over time.
+The marginal utility is defined as the vertical component of the gradient of infected individuals.
+The plot draws on a base population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.}
+	\label{fig:marginal_utility}
+\end{figure}
+
+% subsection population_strategy_contingent_solutions_to_population_dynamics (end)
+
+\subsection{Cost, risk and strategy} % (fold)
+\label{sub:cost_risk_and_strategy}
+
+Any strategy $\sigma$ has a cost $J(\sigma)$, as stated in Section~\ref{sec:introduction}, and the overall societal cost of $n$ individuals $p_{1 \ldots n} \in P$ each adopting, respectively, strategy $\sigma_{1 \ldots n}$, is $\sum_{i=0}^n J(\sigma_i)$.
+But since strategies are limited (one may, at any given time, either engage in social distancing or not, assuming for simplicity's sake that those who do so are entirely successful), for any society-level strategy $\bar{\sigma} (t)$ resulting in a level of distancing described by $\delta (t)$,
+
+\begin{equation}
+	\begin{aligned}
+		\bar{J}(P, t) = \sum_{i=0}^{\delta(P, t) n} J_{\delta} + \sum_{j=0}^{(1-\delta(P, t)) n} J_{\lnot \delta}
+	\end{aligned}
+	\label{eq:j_bar}
+\end{equation}
+
+\noindent where $J_{\delta}$ is the cost of social distancing for discrete unit time and $J_{\lnot \delta}$ is the cost of not distancing for the same unit time.
+The latter of these is not constant, as Equation~\eqref{eq:risk} shows, but a function of a constant cost of infection, $c_i$, and the risk of infection ($r_i$), which is contingent on $I(t)$ and $\delta(t)$.
+Thus, Equation~\eqref{eq:j_bar} can be reformulated (once again, in discrete time) as
+
+\begin{equation}
+	\begin{aligned}
+		\bar{J}(P, t) = \delta(t) n c_d + (1 - \delta(t)) n J_{\lnot \delta}
+	\end{aligned}
+\end{equation}
+
+\noindent which expands to
+
+\begin{equation}
+	\begin{aligned}
+		\bar{J}(P, t) = \delta(t) n c_d + (1 - \delta(t)) n r_i(t) c_i
+	\end{aligned}
+\end{equation}
+
+For a susceptible individual $p_i \in S$, the risk of infection $r_i(t)$ in discrete time is the proportional likelihood of infection, or in other words,
+
+\begin{equation}
+	\begin{aligned}
+		r_i(t) = \frac{\beta S(t) I(t)}{n^2}
+	\end{aligned}
+\end{equation}
+
+While quantification of costs of illness is difficult, quantification of the economic, social and emotional costs of social distancing is even harder to grasp.
+However, we may, for a range of cost fractions $\frac{c_d}{c_i}$, estimate values of $\delta(t)$ that denote optimal strategic spaces.
+Thus,
+
+\begin{equation}
+	\begin{aligned}
+		\frac{c_d}{c_i}(P, t) = \frac{(1 - \delta(t)) R_0 \gamma S(t) I(t)}{\delta(t) n^2}
+	\end{aligned}
+\end{equation}
+
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/cost_fraction}
+	\caption{Cost fraction $\frac{c_d}{c_i}$ of social distancing in a population $P$ of $\delta(P)$ adherence over time, based on a population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
+The contour lines indicate the cost fraction, i.e.
+what fraction of the cost of social distancing $c_d$ the cost of illness $c_i$ must be in order to make not distancing a preferred strategy.}
+	\label{fig:cost_fraction}
+\end{figure}
+
+As numerical estimation of this cost fraction (Figure~\ref{fig:cost_fraction}) shows that for most cases, not distancing can only be a feasible strategy if the cost of distancing very significantly outweighs the cost of illness, often by as much as four orders of magnitude.
+Since this is notwithstanding risk (i.e.
+the comparison needs to be made between the absolute cost of infection and the absolute cost of distancing, rather than the probability/risk-adjusted figures).
+
+% subsection cost_risk_and_strategy (end)
 
 % section methods (end)
+
+\section{Results} % (fold)
+\label{sec:results}
+
+This risk, viewed as a function of $R_0$ and $\delta(P)$ in Figure~\ref{fig:risk_and_compliance}, indicates the somewhat counterintuitive finding that social distancing becomes more,
+
+% section results (end)
 

### 2020-07-23 paper/paper.tex
First draft.

@@
 {
-  pdftitle   = {Marginal dynamics of social distancing in COVID-19},
+  pdftitle   = {Statistical dynamics of SARS-CoV-2 as a differential game},
   pdfauthor  = {Chris von Csefalvay}
@@
 
-\title{Marginal dynamics of social distancing in COVID-19}
+\title{Statistical dynamics of SARS-CoV-2 as a differential game}
 \author{Chris von Csefalvay}
@@
 	\includegraphics[width=\linewidth]{figures/fig1-odes}
-	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000 and a seed population of 1\% infected initially.
+	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000 and a seed population of 0.1\% infected initially.
 For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
@@
 	\includegraphics[width=\linewidth]{figures/fig3-SIR-by-delta}
-	\caption{Contour diagrams of numeric solutions for susceptible, infectious and removed ($S$, $I$ and $R$) compartment sizes for a base population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
+	\caption{Contour diagrams of numeric solutions for susceptible, infectious and removed ($S$, $I$ and $R$) compartment sizes for a base population of 10,000 individuals with a seed population of 0.1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
 As this figure indicates, the 'critical mass' of social distancing takes place in the $\delta$ range of 0 to 0.4, and thus even modest increases in social distancing participation at low levels can make a significant difference in the number of infectious cases.}
@@
 The marginal utility is defined as the vertical component of the gradient of infected individuals.
-The plot draws on a base population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.}
+The plot draws on a base population of 10,000 individuals with a seed population of 0.1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.}
 	\label{fig:marginal_utility}
@@
 	\end{aligned}
+	\label{eq:cost_fraction}
 \end{equation}
@@
 	\includegraphics[width=\linewidth]{figures/cost_fraction}
-	\caption{Cost fraction $\frac{c_d}{c_i}$ of social distancing in a population $P$ of $\delta(P)$ adherence over time, based on a population of 10,000 individuals with a seed population of 1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
+	\caption{Cost fraction $\frac{c_d}{c_i}$ of social distancing in a population $P$ of $\delta(P)$ adherence over time, based on a population of 10,000 individuals with a seed population of 0.1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
 The contour lines indicate the cost fraction, i.e.
@@
 
-This risk, viewed as a function of $R_0$ and $\delta(P)$ in Figure~\ref{fig:risk_and_compliance}, indicates the somewhat counterintuitive finding that social distancing becomes more,
+\subsection{Strategies of social distancing} % (fold)
+\label{sub:strategies_of_social_distancing}
+
+As Figure~\ref{fig:fig3-SIR-by-delta} indicates for a given $R_0$ and $\frac{\beta}{\gamma}$, social distancing can have an overwhelmingly significant effect on the number of infectious individuals in a closed population, dependent on the number of persons in the population already engaged in social distancing.
+This effect is most pronounced early in the epidemic (approx.
+2-3 $\gamma^{-1}$ days), and the effect is most significant where less than half of the population is engaged in social distancing.
+Thus, unlike collective immunity in the case of vaccination, which often necessitates a fairly high level of penetration (typically estimated as $1 - R_0^{-1}$), social distancing can play a meaningful role in particular where much of the population is not engaged in such behaviours.
+This result can meaningfully inform a policy of encouraging individual social distancing early in an outbreak and dispelling perceptions that unless a critical volume of individuals are participating, marginal action is unnecessary.
+
+% subsection strategies_of_social_distancing (end)
+
+\subsection{Marginal utility of social distancing} % (fold)
+\label{sub:marginal_utility_of_social_distancing}
+
+Based on the key epidemiological dynamics data on the SARS-CoV-2 pandemic, the highest marginal effect of social distancing takes place in the same early timeframe of approx.
+2-3 $\gamma^{-1}$ days.
+Unsurprisingly, even without integrating the time-dependent discount factor proposed by Reluga (2010), the numerical solutions indicate that the effect of social distancing is most significant where it is not yet a widely adopted strategy: for SARS-CoV-2, based on an initial population of 10,000 with a seed population of 0.1\% infected, the greatest marginal utility is encountered where less than 20\% of the population is engaged in such behaviour, and the effect is significantly less noticeable once $\delta(P)$ reaches 0.5.
+
+Calculations of marginal utility matter because they can guide public policy in determining what fraction of the population may feasibly be exempted from social distancing.
+Given the need for critical services to continue even in the throes of a pandemic, marginal utility calculations based on empirical data on an outbreak itself, as well as the societal response (as expressed by $\delta(P, t)$), may contribute to more accurate public health measures while limiting the effect of such measures on the economy.
+
+% subsection marginal_utility_of_social_distancing (end)
+
+\subsection{Costs and strategy choices} % (fold)
+\label{sub:costs_and_strategy_choices}
+
+The costs both of social distancing and of failing to do so are notoriously difficult to quantify accurately.
+Indeed, many of these costs are by their very nature not amenable to quantification.
+At the same time, by quantifying the relative ratio of cost of distancing ($c_d$) and cost of illness ($c_i$), we can identify a strategy-associated cost relationship that, given solid information, can assist in societal decision-making as to social distancing.
+As Figure~\ref{fig:cost_fraction} shows, for most cases, the cost of social distancing would have to exceed the cost of illness by at least an order of magnitude to make it a preferable strategy.
+In addition, the computational solution of Equation~\eqref{eq:cost_fraction} shows not only that failure to socially distance may only be a preferable choice if the costs of distancing vastly exceed the costs of illness, but that this remains the case for much of the short term (<90 days).
+
+% subsection costs_and_strategy_choices (end)
 
 % section results (end)
+
+\section{Discussion} % (fold)
+\label{sec:discussion}
+
+Pandemics pose a significant challenge to public health and social decision-making, and the COVID-19 pandemic is by no means an exception.
+Non-pharmaceutical interventions, such as social distancing, play a significant role in the public health arsenal and may play a determinative role in stemming the tide of an infectious disease that is otherwise not amenable to treatment or prophylaxis.
+Thus, until a vaccine or a reliable therapeutic, ideally with prophylactic properties, is found, non-pharmaceutical interventions are the mainstay of public health activity in the face of COVID-19.
+
+This paper discussed a subject that is not devoid of controversy, both in the scientific and in the public realm.
+By their nature, non-pharmaceutical interventions interfere with citizens' day-to-day lives and may have complex economic, social and psychological effects.
+It is therefore important that strategy options are adequately explored from a quantitative perspective.
+It is hoped that in reinforcing the case for social distancing through an analysis of the statistical dynamics that underlie it, this paper can add to the wealth of literature in support of social distancing as an effective and cost-efficient non-pharmaceutical intervention where other tools are unavailable or inappropriate.
+
+% section discussion (end)
+
+\section*{Competing interests} % (fold)
+\label{sec:competing_interests}
+
+The author declares no competing interests.
+
+% section competing_interests (end)
+
+\section*{Supplementary data} % (fold)
+\label{sec:supplementary_data}
+
+All simulations, code and data are available on Github and under the DOI XXXXXXX.
+
+% section supplementary_data (end)
 

### 2020-07-23 paper/paper.tex
23JUL proof version.

@@
 \begin{abstract}
-    Abstract
+    The novel coronavirus SARS-CoV-2 has rapidly emerged as a significant worldwide threat to public health, in particular because -- as is not uncommon with novel pathogens -- there is no effective pharmaceutical treatment or prophylaxis to the viral syndrome it causes.
+In the absence of such specific treatment modalities, the mainstay of public health response rests on non-pharmaceutical interventions (NPIs), such as social distancing.
+This paper contributes to the understanding of social distancing against SARS-CoV-2 by quantitatively analysing the statistical dynamics of disease propagation as a differential game, and estimating the relative costs of distancing versus not distancing, identifying marginal utility of distancing based on known population epidemiological data about SARS-CoV-2 and concluding that unless the costs of distancing vastly exceed the cost of illness per unit time, social distancing remains a dominant strategy.
+These findings can assist in solidly anchoring public health responses based on social distancing within a quantitative framework attesting to their effectiveness.
 \end{abstract}
@@
 This is the case with novel infectious diseases that have no specific treatment and no prophylactic (vaccine) available.
-In the absence of pharmaceutical interventions of proven effectiveness, in particular prophylactically, the main public health response to the emerging pandemic of COVID-19, a viral syndrome caused by the (+)ssRNA virus SARS-CoV-2 (order \emph{Nidovirales}, family \emph{Coronaviridae}, genus \emph{Betacoronavirus}, subgenus \emph{Sarbecovirus}), has rested principally on NPIs, first and foremost social distancing and ancillary steps intended to facilitate that.
+In the absence of pharmaceutical interventions of proven effectiveness, in particular prophylactically, the main public health response to the emerging pandemic of COVID-19, a viral syndrome caused by the (+)ssRNA virus SARS-CoV-2 (order \emph{Nidovirales}, family \emph{Coronaviridae}, genus \emph{Betacoronavirus}, subgenus \emph{Sarbecovirus}), has rested principally on NPIs.\cite{mccoy2020can,lai2020effect,flaxman2020report,ferguson2020report} At their core, all NPIs share a quintessential relationship to social distancing, either through directly encouraging social distancing, limiting encounters by reducing public facilities for such encounters that may transmit the pathogen ('lockdowns'), reducing gatherings and social interactions that carry such risk ('large-gathering bans'), suspending economic activities that inherently carry the risk of social interactions or modifying the framework of such activities to reduce such interactions (e.g.
+transitioning to remote work).
 
-In general, any course of conduct that reduces the encounter rate between an individual and other individuals can be considered a form of social distancing.
-This may be brought about through limiting public facilities for such encounters ('lockdowns'), through limiting individual gatherings by size ('large-gathering bans') and through encouraging individual social distancing.
-From the perspective of game theory, social distancing can be viewed as a non-cooperative game of a population $P_{1 \ldot n}$ of size $n$, where at any given time $t \in [t_0, t_{e}]$, the strategy adopted by $p_i$ is denoted as $\sigma (p_i, t)$.
-For the sake of simplicity, we assume that distancing is either not exercised at all or perfectly exercised, i.e.
-$\sigma (p_i, t) \in \{ 0, 1 \}$.
-Then, for the entire population, the overall strategy can be described as
+From the perspective of game theory, social distancing can be viewed as a non-cooperative game of a population $P_{1 \ldots n}$ of size $n$, where at any given time $t \in [t_0, t_f]$, each player $p_i$ adopts the strategy $\sigma(p_i, t)$.
+For the sake of simplicity, we will assume two fundamental properties of social distancing conduct:
+
+\begin{enumerate}
+	\item There exist two strategy choices, $\delta$ and $\lnot \delta$ (distancing and not distancing, respectively).
+At any given time, any agent $p_i \in P$ can opt exclusively for one of these two options, i.e.
+the strategy set $\mathcal{S}(p_i)$ is the set $\{\delta, \lnot \delta\}$.
+	\item Where an agent $p_i$ makes a policy choice at time $t$, $\sigma(p_i, t) \in [\delta, \lnot \delta]$, they implement it perfectly, i.e.
+the efficiency of every actor $p_i \in P$ in implementing their strategy $\sigma(p_i, t)$ is equal.
+\end{enumerate}
+
+Then, for the entire population $P$, the overall strategy $\bar{\sigma}(P, t)$ can be described as the sum of all strategies $\sigma(p_i, t) | p_i \in P$.
+We can then assign a value $\hat{\sigma}$ to each strategy, whereby the strategy of distancing, $\delta$, is assigned the value of 1 and the strategy of not distancing, $\lnot \delta$, the value of 0.
+
+For any given discrete time $t \in [t_0, t_f]$, where $t_f$ would be an endpoint (such as eradication, elimination, natural extinction of the pathogen, the availability of a vaccine or a combination thereof), we may then define two disjoint sets, $P_{\delta}(t)$ and $P_{\lnot \delta} (t)$, where
 
 \begin{equation}
-	\bar{\sigma}(P, t) = \frac{\displaystyle \sum_{i = 0}^n \sigma(p_i, t)}{n}
-	\label{eq:overall_strategy}
-\end{equation}
-
-\noindent and for the entire time period, from $t_0$ to the endpoint $t_e$ (which may be eradication, elimination, natural extinction of the pathogen, the availability of a vaccine or a combination thereof), for discrete time $t$,
-
-\begin{equation}
-	\bar{\bar{\sigma}}(P) = \frac{\displaystyle \sum_{i = 0}^n \displaystyle \sum_{j = 0}^{t_e - t_0} \sigma(p_i, j)}{n (t_e - t_0)}
-	\label{eq:overall_strategy_over_time}
-\end{equation}
-
-However, with each course of conduct, there is associated a cost $J(\sigma)$.
-For simplicity's sake, let us consider these costs to be governed by the following three precepts given a pathogen with fixed characteristics ($R_0$, transmission potential, risks \&c.).
-Let $\bar{\delta}(P)$ equal the proportion of persons $p \in P$ adopting strategy $\sigma_{\delta}$.
-It then holds that:
-
-\begin{enumerate}
-	\item A person $p_i \in P$ opting for strategy $\delta$ (social distancing) will incur $c_d$, the immediate costs of distancing.
-These may be social (lessened social interaction), psychological (lessened access to support systems), economic (lower access to facilities to earn) or simple matters of convenience (access to amenities).
-While $c_d$ is somewhat dependent on $\bar{\bar{(P)}}$ (thus not distancing does not yield a benefit to a lone social distancer in terms of access to amenities if all of the latter are closed), it can be assumed to be largely constant.
-	\item Compared to a person opting for strategy $\delta$, a person $p_i \in P$ opting for strategy $\lnot \delta$ will incur a relative additional risk $f_r(\bar{\delta}(P), t)$, i.e.
-contingent on the population's behaviour at time $t$.
-	\item Finally, a person $p_i \in P$ will regardless of their individual choice (at least at non-trivial population sizes) gain a protective benefit, which is a function of the population's adherence to social distancing.
-We will denote this term as $f_c(\bar{\delta}(P), t)$.
-\end{enumerate}
-
-Then, assigning the variable $c_s$ to represent the cost of illness (economic loss, medical costs, long-term health risks), we can express the cost functions of each strategy, $\sigma_{\delta}$ (social distancing) and $\sigma_{\lnot \delta}$ (no social distancing), as
-
-\begin{equation}
-	J_{\sigma_{\delta}}(p_i, t) = c_d - f_c(\bar{\delta}(P), t)
+	\forall p_j \in P_{\delta}(t), \ \ \sigma(p_j, t) = \delta
 \end{equation}
@@
 \begin{equation}
-	J_{\sigma_{\lnot \delta}}(p_i, t) = f_r(\bar{\delta}(P), t) c_s - f_c(\bar{\delta}(P), t)
-	\label{eq:risk}
+	\forall p_k \in P_{\lnot \delta}(t), \ \ \sigma(p_k, t) = \lnot \delta
 \end{equation}
 
-\noindent And consequently, it holds for the population level cost $\bar{J}$ at time $t$ that
+\noindent so that $|P_{\delta}(t)| + |P_{\lnot \delta}(t)| = |P| = n$.
+Then, for any $t$ in discrete time $t \in [t_0, t_f]$, we define the overall factor of social distancing for a population $P$ at time $t$, $\delta(P, t)$
 
 \begin{equation}
-	\bar{J}(P, t) = \frac{\displaystyle \sum_{i=0}^{\delta(P) n} J_{\sigma_{\delta}}(p_i, t) + \displaystyle \sum_{j=0}^{1-\delta(P) n} J_{\sigma_{\lnot \delta}}(p_j, t)}{n}
-	\label{eq:cost_eqn}
+	\begin{aligned}
+		\delta(P, t) = \frac{P_{\delta}(t)}{n} = 1 - \frac{P_{\lnot \delta}(t)}{n}
+	\end{aligned}
 \end{equation}
 
-\noindent and for a population over time $t_0$ to $t_e$,
+\noindent which can also be expressed as a function of strategies $\sigma(p_{1 \ldots n})$ as
 
 \begin{equation}
-	\bar{\bar{J}}(P) = \int_{t_0}^{t_e} \bar{J}(P, t) dt
+	\begin{aligned}
+		\delta(P, t) = \frac{\sum_{i=1}^n \hat{\sigma}(p_i, t)}{n}
+	\end{aligned}
+	\label{eq:delta_value}
 \end{equation}
 
-As Reluga (2010) notes, the effect of social distancing diminishes with the increase in participants, the phenomenon of diminishing returns.\cite{reluga2010game} Thus, as the number of individuals in $P$ opting for social distancing increases, the marginal increase by another social distancer diminishes.
-This can be represented by a discount factor $h$, to yield
+In other words, the overall factor of social distancing at time $t$ is the average of the value assigned to each agent's strategy over the entire population $P$.
+For simplicity's sake, we will assume that the decision process takes the shape of a continuous and simultaneous game in discrete time, and agents can instantly switch strategies with no cost (other than the cost of the strategy itself).
+It then holds that
+
+\begin{enumerate}
+	\item A person $p_i \in P$ opting for strategy $\sigma_{\delta}$ (social distancing) will incur $c_d$, the immediate costs of distancing.
+These may be social (lessened social interaction), psychological (lessened access to support systems), economic (lower access to facilities to earn) or simple matters of convenience (access to amenities).
+While $c_d$ is somewhat dependent on $\delta(P, t)$ (thus not distancing does not yield a benefit to a lone social distancer in terms of access to amenities if all of the latter are closed), it can be assumed to be largely constant.
+	\item Compared to a person opting for strategy $\sigma_{\delta}$, a person $p_i \in P$ opting for strategy $\sigma_{\lnot \delta}$ will incur a relative additional risk $r_i(P, t) c_i$, where $c_i$ denotes the constant cost of illness and $r_i(P, t)$ is the risk of contracting illness when not distancing, a function of $\delta(P, t)$.
+\end{enumerate}
+
+This allows us to identify the cost function $J(\sigma, t)$ for any $\sigma \in \{\delta, \lnot \delta\}$ at time $t$ for each individual $p_i \in P$ as
 
 \begin{equation}
-	\bar{J}(P) = \int_{t_0}^{t_e} e^{-ht} \frac{\delta(P) J_{\sigma_{\delta}}(P, t) + (1 - \delta(P)) J_{\sigma_{\lnot \delta}}(P, t)}{n}
+	\begin{aligned}
+		J(\sigma_{\delta}, t) = c_d 									\\
+		J(\sigma_{\lnot \delta}, t) = r_i(P, t) c_i
+	\end{aligned}
+	\label{eq:pure_time_dependent_costs}
 \end{equation}
 
-However, because individual decisions affect the overall gain (due to the component dependent on $\delta(P, t)$), our principal concern is not with individual action but with analysis of such strategies on a population level.
-This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution} and, since its publication in 1927, widely adapted and adopted.\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization}
+Then, for every social strategy $\bar{\sigma}(P, t)$ associated with the social distancing value $\delta(P, t)$ in the way described in Equation~\eqref{eq:delta_value}, the overall social cost of $\bar{\sigma}(P, t)$ can be conceptualised as
+
+\begin{equation}
+	\begin{aligned}
+		\bar{J}(P, t) = n \Big( \delta(P, t) c_d + (1-\delta(P, t)) r_i(P, t) c_i \Big)
+	\end{aligned}
+	\label{eq:social_time_dependent_costs}
+\end{equation}
+
+For the entire time space $[t_0, t_f]$, the total social cost is then
+
+\begin{equation}
+	\bar{\bar{J}}(P) = \sum_{t_i = 0}^{t_f - t_0} J(P, t_i) = n \sum_{t_i = 0}^{t_f - t_0} \Big( \delta(P, t_i) c_d + (1 - \delta(P, t_i)) r_i(P, t_i) c_i \Big)
+\end{equation}
+
+Our principal concern in this paper is not with individual action but with analysis of decision strategies on a population level.
+This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution} and, since its publication in 1927, widely adapted and adopted,\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization} and identify within the statistical dynamics of that differential game the equilibria that govern ideal societal decision-making.
 
@@
 	\begin{aligned}
-		\frac{dS}{dt} = - \frac{\beta S I}{n} 								\\
-		\frac{dI}{dt} = \frac{\beta S I}{n} - \gamma I 						\\
-		\frac{dR}{dt} = \gamma I
+		\frac{dS}{dt} &= - \frac{\beta S I}{n} 								\\
+		\frac{dI}{dt} &= \frac{\beta S I}{n} - \gamma I 					\\
+		\frac{dR}{dt} &= \gamma I
 	\end{aligned}
@@
 	\includegraphics[width=\linewidth]{figures/fig1-odes}
-	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma$ between $1/4.6$ and $1/12.4$ over a base population of 10,000 and a seed population of 0.1\% infected initially.
+	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma^{-1}$ between $4.6$ and $12.4$ over a base population of 10,000 and a seed population of 0.1\% infected initially.
 For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
@@
 The fraction $\frac{\beta}{\gamma}$ equals the basic reproduction number, $R_0$.
-For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the invers of the number of days of illness, which can be approximated as $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
+For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the inverse of the average number of days of illness ($\gamma^{-1}$, sometimes also described as $\tau$), which can be approximated as $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
 Even in the absence of firm evidence as to whether SARS-CoV-2 infection followed by recovery would engender lifelong immunity or not,\cite{roy2020covid,ota2020will,lin2020duration} it can be assumed in the short term -- based on evidence from MERS-CoV and SARS-CoV -- that in the short term, survivors remain immune,\cite{prompetchara2020immune} and consequently the number of individuals in $R$ does not decrease.
@@
 	\begin{aligned}
-		\frac{dS}{dt} = - \frac{\beta S I - \delta \beta S I}{n} 			\\
-		\frac{dI}{dt} = \frac{\beta S I - \delta \beta S I}{n} - \gamma I	\\
-		\frac{dR}{dt} = \gamma I
+		\frac{dS}{dt} &= - \frac{\beta S I - \delta \beta S I}{n} 			\\
+		\frac{dI}{dt} &= \frac{\beta S I - \delta \beta S I}{n} - \gamma I	\\
+		\frac{dR}{dt} &= \gamma I
 	\end{aligned}
@@
 While quantification of costs of illness is difficult, quantification of the economic, social and emotional costs of social distancing is even harder to grasp.
-However, we may, for a range of cost fractions $\frac{c_d}{c_i}$, estimate values of $\delta(t)$ that denote optimal strategic spaces.
-Thus,
+However, we may, values of $\delta(P)$, calculate cost fractions $\phi(P, t) = \frac{c_d}{c_i}(P, t)$, where
 
@@
 	\begin{aligned}
-		\frac{c_d}{c_i}(P, t) = \frac{(1 - \delta(t)) R_0 \gamma S(t) I(t)}{\delta(t) n^2}
+		\phi(P, t) = \frac{c_d}{c_i}(P, t) = \frac{(1 - \delta(t)) R_0 \gamma S(t) I(t)}{\delta(t) n^2}
 	\end{aligned}
@@
 
-\begin{figure}
-	\includegraphics[width=\linewidth]{figures/cost_fraction}
-	\caption{Cost fraction $\frac{c_d}{c_i}$ of social distancing in a population $P$ of $\delta(P)$ adherence over time, based on a population of 10,000 individuals with a seed population of 0.1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
-The contour lines indicate the cost fraction, i.e.
-what fraction of the cost of social distancing $c_d$ the cost of illness $c_i$ must be in order to make not distancing a preferred strategy.}
-	\label{fig:cost_fraction}
-\end{figure}
-
-As numerical estimation of this cost fraction (Figure~\ref{fig:cost_fraction}) shows that for most cases, not distancing can only be a feasible strategy if the cost of distancing very significantly outweighs the cost of illness, often by as much as four orders of magnitude.
-Since this is notwithstanding risk (i.e.
-the comparison needs to be made between the absolute cost of infection and the absolute cost of distancing, rather than the probability/risk-adjusted figures).
+This cost fraction indicates the relative disequilibrium between the cost of distancing and the cost of illness when adjusted for risk -- in other words, for any given value of $\delta(P, t)$ at $t$, the cost fraction indicates an inflection point.
+As long as the cost of distancing is less than $\phi c_i$, social distancing at $\delta(P, t)$ or above is the optimum strategy.
+As numerical estimation of this cost fraction (Figure~\ref{fig:cost_fraction}) shows, social distancing is almost always the preferred strategy at equal cost (black line).
+The contour lines in Figure~\ref{fig:cost_fraction} indicate what ratio the cost of distancing has to be to the cost of illness at a given fraction $\delta(P, t)$ of distancing to make distancing no longer an optimal strategy.
+So, for instance, a $\phi$ of 0.05 denotes the isorisk curve over $t$ where, at given $\delta(P, t)$, distancing is the preferable solution as long as its costs are 0.05 times that of illness or less.
 
@@
 
+\begin{figure}
+	\includegraphics[width=\linewidth]{figures/cost_fraction}
+	\caption{Cost fraction $\phi = \frac{c_d}{c_i}$ of social distancing in a population $P$ of $\delta(P, t)$ adherence over time, based on a population of 10,000 individuals with a seed population of 0.1\% infected, under the assumption of an $R_0$ of 2.67 and $\gamma$ of $\frac{1}{8.5}$.
+The contour lines indicate the cost fraction, i.e.
+what fraction of the cost of social distancing $c_d$ the cost of illness $c_i$ must be in order to make not distancing a preferred strategy.}
+	\label{fig:cost_fraction}
+\end{figure}
+
 The costs both of social distancing and of failing to do so are notoriously difficult to quantify accurately.
@@
 In addition, the computational solution of Equation~\eqref{eq:cost_fraction} shows not only that failure to socially distance may only be a preferable choice if the costs of distancing vastly exceed the costs of illness, but that this remains the case for much of the short term (<90 days).
+
+Estimates of direct medical costs of COVID-19 are difficult, but at least one figure puts the median cost of a symptomatic infection at US\$3,045 in direct costs alone,\cite{bartsch2020potential} typically compounded by loss of earnings, long-term physical harm, reduction in life expectancy and quality of life and, in severe cases, the risk of mortality.
+The costs of social distancing are much less amenable to quantification, as these costs are primarily governed by indirect factors and intangibles, such as the cost of deferred medical treatment, second-order effects of the psychological burden inherent in decreased social interaction and the cost of lost revenue.
+While quantification, thus, of both the cost of distancing and cost of illness remains an outstanding subject of research, the cost fraction calculations can assist in reasoning about the best social strategy in view of these factors once ascertained.
 

### 2020-07-23 paper/paper.tex
Final draft.

@@
 % Copyright (c) Chris von Csefalvay, 2020.
-\documentclass{article}
+\documentclass[12pt]{article}
 \usepackage{amsmath}
@@
 {
-  pdftitle   = {Statistical dynamics of SARS-CoV-2 as a differential game},
+  pdftitle   = {Statistical dynamics of social distancing in SARS-CoV-2 as a differential game},
   pdfauthor  = {Chris von Csefalvay}
@@
 
-\title{Statistical dynamics of SARS-CoV-2 as a differential game}
-\author{Chris von Csefalvay}
+\title{Statistical dynamics of social distancing in SARS-CoV-2 as a differential game}
+\author{Chris von Csefalvay\thanks{Starschema Inc., Arlington, VA. Correspondence: \texttt{csefalvayk@starschema.net}.}}
 
@@
 \begin{abstract}
-    The novel coronavirus SARS-CoV-2 has rapidly emerged as a significant worldwide threat to public health, in particular because -- as is not uncommon with novel pathogens -- there is no effective pharmaceutical treatment or prophylaxis to the viral syndrome it causes.
+    The novel coronavirus SARS-CoV-2 has rapidly emerged as a significant threat to global public health, in particular because -- as is not uncommon with novel pathogens -- there is no effective pharmaceutical treatment or prophylaxis to the viral syndrome it causes.
 In the absence of such specific treatment modalities, the mainstay of public health response rests on non-pharmaceutical interventions (NPIs), such as social distancing.
@@
 This is the case with novel infectious diseases that have no specific treatment and no prophylactic (vaccine) available.
-In the absence of pharmaceutical interventions of proven effectiveness, in particular prophylactically, the main public health response to the emerging pandemic of COVID-19, a viral syndrome caused by the (+)ssRNA virus SARS-CoV-2 (order \emph{Nidovirales}, family \emph{Coronaviridae}, genus \emph{Betacoronavirus}, subgenus \emph{Sarbecovirus}), has rested principally on NPIs.\cite{mccoy2020can,lai2020effect,flaxman2020report,ferguson2020report} At their core, all NPIs share a quintessential relationship to social distancing, either through directly encouraging social distancing, limiting encounters by reducing public facilities for such encounters that may transmit the pathogen ('lockdowns'), reducing gatherings and social interactions that carry such risk ('large-gathering bans'), suspending economic activities that inherently carry the risk of social interactions or modifying the framework of such activities to reduce such interactions (e.g.
+In the absence of pharmaceutical interventions of proven effectiveness, in particular prophylactically, the main public health response to the emerging pandemic of COVID-19, a viral syndrome caused by the (+)ssRNA virus SARS-CoV-2 (order \emph{Nidovirales}, family \emph{Coronaviridae}, genus \emph{Betacoronavirus}, subgenus \emph{Sarbecovirus}), has rested principally on NPIs.\cite{McCoy_2020,Lai_2020,flaxman2020report,ferguson2020report} At their core, all NPIs share a quintessential relationship to social distancing, either through directly encouraging social distancing, limiting transmission potential by reducing public facilities for such encounters that may transmit the pathogen ('lockdowns'), reducing gatherings and social interactions that carry such risk ('large-gathering bans'), suspending economic activities that inherently carry the risk of social interactions or modifying the framework of activities to reduce such interactions (e.g.
 transitioning to remote work).
@@
 From the perspective of game theory, social distancing can be viewed as a non-cooperative game of a population $P_{1 \ldots n}$ of size $n$, where at any given time $t \in [t_0, t_f]$, each player $p_i$ adopts the strategy $\sigma(p_i, t)$.
-For the sake of simplicity, we will assume two fundamental properties of social distancing conduct:
+For the sake of simplicity, we will assume two fundamental axioms about social distancing behaviours:
 
@@
 the strategy set $\mathcal{S}(p_i)$ is the set $\{\delta, \lnot \delta\}$.
-	\item Where an agent $p_i$ makes a policy choice at time $t$, $\sigma(p_i, t) \in [\delta, \lnot \delta]$, they implement it perfectly, i.e.
-the efficiency of every actor $p_i \in P$ in implementing their strategy $\sigma(p_i, t)$ is equal.
+	\item Where a player $p_i$ makes a strategy choice $\sigma(p_i, t) \in \mathcal{S}(p_i)$ at time $t$, they implement it perfectly, i.e.
+the efficiency of every player $p_i \in P$ in implementing their strategy $\sigma(p_i, t)$ is equal.
 \end{enumerate}
 
-Then, for the entire population $P$, the overall strategy $\bar{\sigma}(P, t)$ can be described as the sum of all strategies $\sigma(p_i, t) | p_i \in P$.
+Then, for the entire population $P$, the aggregate population level strategy $\bar{\sigma}(P, t)$ can be described as the sum of all strategies $\sigma(p_i, t) | p_i \in P$.
 We can then assign a value $\hat{\sigma}$ to each strategy, whereby the strategy of distancing, $\delta$, is assigned the value of 1 and the strategy of not distancing, $\lnot \delta$, the value of 0.
@@
 \begin{equation}
-	\forall p_j \in P_{\delta}(t), \ \ \sigma(p_j, t) = \delta
-\end{equation}
-
-\noindent and
-
-\begin{equation}
-	\forall p_k \in P_{\lnot \delta}(t), \ \ \sigma(p_k, t) = \lnot \delta
+	\begin{aligned}
+		\forall p_j \in P_{\delta}(t), & \ \ \sigma(p_j, t) = \delta						\\
+		\forall p_k \in P_{\lnot \delta}(t), & \ \ \sigma(p_k, t) = \lnot \delta		
+	\end{aligned}
 \end{equation}
@@
 
-In other words, the overall factor of social distancing at time $t$ is the average of the value assigned to each agent's strategy over the entire population $P$.
+In other words, the overall factor of social distancing at time $t$, $\delta(P, t)$ is the average of the value the function $\hat{\sigma}$ assigns to each player's strategy over the entire population $P$.
 For simplicity's sake, we will assume that the decision process takes the shape of a continuous and simultaneous game in discrete time, and agents can instantly switch strategies with no cost (other than the cost of the strategy itself).
@@
 \begin{enumerate}
-	\item A person $p_i \in P$ opting for strategy $\sigma_{\delta}$ (social distancing) will incur $c_d$, the immediate costs of distancing.
+	\item A player $p_i \in P$ opting for strategy $\sigma_{\delta}$ (social distancing) will incur $c_d$, the immediate costs of distancing.
 These may be social (lessened social interaction), psychological (lessened access to support systems), economic (lower access to facilities to earn) or simple matters of convenience (access to amenities).
-While $c_d$ is somewhat dependent on $\delta(P, t)$ (thus not distancing does not yield a benefit to a lone social distancer in terms of access to amenities if all of the latter are closed), it can be assumed to be largely constant.
-	\item Compared to a person opting for strategy $\sigma_{\delta}$, a person $p_i \in P$ opting for strategy $\sigma_{\lnot \delta}$ will incur a relative additional risk $r_i(P, t) c_i$, where $c_i$ denotes the constant cost of illness and $r_i(P, t)$ is the risk of contracting illness when not distancing, a function of $\delta(P, t)$.
+While $c_d$ is somewhat dependent on $\delta(P, t)$ (thus not distancing does not yield a benefit to a lone player in terms of access to amenities if all of these closed due to widespread social distancing), it can be assumed to be largely constant.
+	\item Compared to a person opting for strategy $\sigma_{\delta}$, a person $p_i \in P$ opting for strategy $\sigma_{\lnot \delta}$ will not incur the fixed cost $c_d$, but will instead incur a relative additional cost $r_i(P, t) c_i$, where $c_i$ denotes the constant cost of illness and $r_i(P, t)$ is the risk of contracting illness when not distancing, a function of $\delta(P, t)$.
 \end{enumerate}
@@
 	\begin{aligned}
-		J(\sigma_{\delta}, t) = c_d 									\\
-		J(\sigma_{\lnot \delta}, t) = r_i(P, t) c_i
+		J(\sigma_{\delta}, t) & = c_d 									\\
+		J(\sigma_{\lnot \delta}, t) & = r_i(P, t) c_i
 	\end{aligned}
@@
 
-Then, for every social strategy $\bar{\sigma}(P, t)$ associated with the social distancing value $\delta(P, t)$ in the way described in Equation~\eqref{eq:delta_value}, the overall social cost of $\bar{\sigma}(P, t)$ can be conceptualised as
+Then, for every population level aggregate strategy $\bar{\sigma}(P, t)$ associated with the social distancing value $\delta(P, t)$ in the way described in Equation~\eqref{eq:delta_value}, the overall social cost of $\bar{\sigma}(P, t)$ can be conceptualised as
 
@@
 
-Our principal concern in this paper is not with individual action but with analysis of decision strategies on a population level.
-This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution} and, since its publication in 1927, widely adapted and adopted,\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization} and identify within the statistical dynamics of that differential game the equilibria that govern ideal societal decision-making.
+The principal concern of this paper is not with individual action but with analysis of decision strategies on a population level.
+This paper will in the following conceptualise infectious disease in a population as a differential game over a differential equation form of the compartmental model first described by Kermack and McKendrick\cite{kermack1927contribution}.
+This model has been widely adapted and adopted since its publication in 1927,\cite{vstvepan2007kermack,roberts1999kermack,capasso1978generalization} and building on it, we will go on to identify within the statistical dynamics of that differential game the equilibria that govern ideal societal decision-making.
 
@@
 
-Given a population of $n$ under the assumption that reinfection is impossible (as, e.g., in the case of measles) or rare (as is the case for SARS-CoV-2\cite{edridge2020human,deng2020primary,bao2020reinfection}), and neglecting for the time being the vital dynamics (birth, unrelated death, migration) of the population, the dynamics of any population can be modelled as a system of ordinary differential equations
+Given a population of $n$ under the assumption that reinfection is sufficiently rare so as to be ignored at a population level, as appears to be the case for SARS-CoV-2,\cite{edridge2020human,deng2020primary,Bao:2020aa}), the dynamics of subpopulations with respect to infection can be modelled as a system of ordinary differential equations
 
@@
 \begin{figure}
-	\includegraphics[width=\linewidth]{figures/fig1-odes}
+	\includegraphics[width=0.8\linewidth]{figures/fig1-odes}
+	\centering
 	\caption{Some quantitative solutions for the SIR model's population dynamics over values of $R_0$ between 1.5 and 6.5, and values of $\gamma^{-1}$ between $4.6$ and $12.4$ over a base population of 10,000 and a seed population of 0.1\% infected initially.
-For each plot, $\beta$ is inferred from an $R_0$ value of 2.67, based on Liu et al.
-(2020),\cite{liu2020reproductive} and the $\gamma$ parameter.
-The susceptible population is displayed in blue, while infected/infectious cases are marked in red and recovered cases in green.}
+For each plot, $\beta$ is inferred from $R_0$ and $\gamma$ using Equation~\eqref{eq:r0_equation}.
+The susceptible population is displayed in blue, while infected/infectious cases are marked in red and removed cases in green.}
 	\label{fig:ode_solutions}
@@
 
-\noindent under the assumption of $S + I + R = 0$, where $S$ represents susceptible individuals, $I$ represents infected/infectious individuals and $R$ accounts for removed individuals (mortality and recovery to immunity).
+\noindent under the assumption of a closed, static population, i.e.
+neglecting for the time being the vital dynamics (birth, unrelated death, migration) of the population.
+Thus, $S + I + R = n$, where $S$ represents susceptible individuals, $I$ represents infected/infectious individuals and $R$ accounts for removed individuals (mortality and recovery to immunity).
+In addition, due to the closed population assumption,
+
+\begin{equation}
+	\frac{dS}{dt} + \frac{dI}{dt} + \frac{dR}{dt} = 0
+\end{equation}
+
+Furthermore, the factors $\beta$ and $\gamma$ in Equation~\eqref{eq:sir_equation} relate to each other as
+
+\begin{equation}
+	R_0 = \frac{\beta}{\gamma} | \gamma > 0
+	\label{eq:r0_equation}
+\end{equation}
+
 The fraction $\frac{\beta}{\gamma}$ equals the basic reproduction number, $R_0$.
-For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the inverse of the average number of days of illness ($\gamma^{-1}$, sometimes also described as $\tau$), which can be approximated as $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$ under the assumption of an $R_0$ of 2.67.
-Even in the absence of firm evidence as to whether SARS-CoV-2 infection followed by recovery would engender lifelong immunity or not,\cite{roy2020covid,ota2020will,lin2020duration} it can be assumed in the short term -- based on evidence from MERS-CoV and SARS-CoV -- that in the short term, survivors remain immune,\cite{prompetchara2020immune} and consequently the number of individuals in $R$ does not decrease.
-Numerical solutions to this system of differential equations have been calculated using \texttt{odepack} via \texttt{SciPy 1.5.1}\cite{virtanen2020scipy} on Python 3.6, and are presented as Figure~\ref{fig:ode_solutions} for a range of values of $R_0$ and $\gamma$.
+For SARS-CoV-2, estimates of $R_0$ range from 1.4 to 6.49, with studies that relied on statistical estimation of $R_0$ ranging from 2.20 to 3.58, with an average of 2.67\cite{liu2020reproductive} $\gamma$, on the other hand, can be estimated as the inverse of the average number of days of illness ($\gamma^{-1}$, sometimes also described as $\tau$).
+This value $\tau$ has been identified by studies of the initial infection dynamics of SARS-CoV-2 to be approximately $8.5 \pm 3.9$ days.\cite{pan2020clinical,liu2020risk} Even in the absence of firm evidence as to whether SARS-CoV-2 infection followed by recovery would engender lifelong immunity or not,\cite{roy2020covid,ota2020will,lin2020duration} it can be assumed in the short term -- based on evidence from MERS-CoV and SARS-CoV -- that at least in the immediate aftermath of disease and recovery, survivors remain immune,\cite{prompetchara2020immune} and consequently $\frac{dR}{dt} \geq 0$ for any $t \in [t_0, t_f]$, i.e.
+the number of removed individuals ($R$) is strictly monotonously increasing over time.
+The inverse is true, for the same reasons, for $\frac{dS}{dt}$ and the number of susceptible individuals ($S$).
+Numerical solutions to this system of differential equations have been calculated using \texttt{odepack} via \texttt{SciPy 1.5.1}\cite{virtanen2020scipy} on Python 3.6, and are described in Figure~\ref{fig:ode_solutions} describes some analytical solutions for the differential equations of Equation~\eqref{eq:sir_equation} over a range of plausible values of $R_0$ and $\gamma$, with $\beta$ inferred from $\gamma$ and $R_0$ through the relationship described in Equation~\eqref{eq:r0_equation}.
 
@@
 
-Given a population that then adopts a sum strategy $\bar{\sigma}(P, t)$ at time $t$ that results in $\delta(P, t)$ adherence (or in Reluga's terms, investment) to social distancing, the flow from $S$ to $I$ is reduced by a corresponding factor.
-This allows us to rewrite Equation~\eqref{eq:sir_equation} so that for a society level strategy $\bar{\sigma}$ yielding $\delta$, the populations can be characterised as
+Given a population that then adopts an aggregate strategy $\bar{\sigma}(P, t)$ at time $t$ that results in $\delta(P, t)$ adherence (or in Reluga's terms, investment\cite{reluga2010game}) to social distancing, the flow from $S$ to $I$ is reduced by a corresponding factor.
+This allows us to rewrite Equation~\eqref{eq:sir_equation} so that for an aggregate strategy $\bar{\sigma}(P, t)$ yielding $\delta$, the populations can be characterised as
 
@@
 the partial derivative of $I(P, t)$ over $\delta(P, t)$.
-Thus, for a population-level strategy $\bar{\sigma}(P, t)$ associated with a compliance rate (i.e.
-the fraction of persons $p_i \in P$ engaged in social distancing at time $t$) of $\delta(P, t)$, there exists a function $\hat{U}(P, t)$ given $\gamma$ and $R_0$ that indicates the marginal utility at any given value of $\delta(P, t)$.
+Thus, for a population-level strategy $\bar{\sigma}(P, t)$ associated with $\delta(P, t)$, there exists a marginal utility function $\hat{U}(P, t)$ given $\gamma$ and $R_0$ that indicates the marginal utility at any given value of $\delta(P, t)$.
 This, too, can be numerically ascertained, and is shown on Figure~\ref{fig:marginal_utility}.
@@
 
-Any strategy $\sigma$ has a cost $J(\sigma)$, as stated in Section~\ref{sec:introduction}, and the overall societal cost of $n$ individuals $p_{1 \ldots n} \in P$ each adopting, respectively, strategy $\sigma_{1 \ldots n}$, is $\sum_{i=0}^n J(\sigma_i)$.
-But since strategies are limited (one may, at any given time, either engage in social distancing or not, assuming for simplicity's sake that those who do so are entirely successful), for any society-level strategy $\bar{\sigma} (t)$ resulting in a level of distancing described by $\delta (t)$,
+Any strategy $\sigma$ has a cost $J(\sigma, t)$, as stated in Section~\ref{sec:introduction}, and the aggregate cost of $n$ individuals $p_{1 \ldots n} \in P$ each adopting, respectively, strategy $\sigma_{1 \ldots n}$, is $\sum_{i=0}^n J(\sigma_i, t)$.
+But since strategies are limited (one may, at any given time, either engage in social distancing or not, assuming for simplicity's sake that those who do so are entirely successful), for any aggregate strategy $\bar{\sigma} (P, t)$ resulting in a level of distancing described by $\delta (P, t)$,
 
@@
 \noindent where $J_{\delta}$ is the cost of social distancing for discrete unit time and $J_{\lnot \delta}$ is the cost of not distancing for the same unit time.
-The latter of these is not constant, as Equation~\eqref{eq:risk} shows, but a function of a constant cost of infection, $c_i$, and the risk of infection ($r_i$), which is contingent on $I(t)$ and $\delta(t)$.
+The latter of these is not constant, as Equation~\eqref{eq:pure_time_dependent_costs} shows, but a function of a constant cost of infection, $c_i$, and the risk of infection ($r_i$), which in turn is contingent on $I(t)$ and $\delta(t)$.
 Thus, Equation~\eqref{eq:j_bar} can be reformulated (once again, in discrete time) as
@@
 	\begin{aligned}
-		\bar{J}(P, t) = \delta(t) n c_d + (1 - \delta(t)) n J_{\lnot \delta}
+		\bar{J}(P, t) = \delta(P, t) n c_d + (1 - \delta(P, t)) n J_{\lnot \delta}
 	\end{aligned}
@@
 	\begin{aligned}
-		\bar{J}(P, t) = \delta(t) n c_d + (1 - \delta(t)) n r_i(t) c_i
+		\bar{J}(P, t) = \delta(P, t) n c_d + (1 - \delta(P, t)) n r_i(t) c_i
 	\end{aligned}
@@
 
-While quantification of costs of illness is difficult, quantification of the economic, social and emotional costs of social distancing is even harder to grasp.
-However, we may, values of $\delta(P)$, calculate cost fractions $\phi(P, t) = \frac{c_d}{c_i}(P, t)$, where
+While quantification of costs of illness is difficult, quantification of the economic, social and emotional costs of social distancing is an even more complex task.
+However, we may, for values of $\delta(P, t)$, calculate cost fractions $\phi(P, t) = \frac{c_d}{c_i}(P, t)$, where
 
@@
 As numerical estimation of this cost fraction (Figure~\ref{fig:cost_fraction}) shows, social distancing is almost always the preferred strategy at equal cost (black line).
-The contour lines in Figure~\ref{fig:cost_fraction} indicate what ratio the cost of distancing has to be to the cost of illness at a given fraction $\delta(P, t)$ of distancing to make distancing no longer an optimal strategy.
-So, for instance, a $\phi$ of 0.05 denotes the isorisk curve over $t$ where, at given $\delta(P, t)$, distancing is the preferable solution as long as its costs are 0.05 times that of illness or less.
+The contour lines in Figure~\ref{fig:cost_fraction} indicate what ratio the cost of distancing has to be to the cost of illness at a given value of $\delta(P, t)$ to make distancing no longer an optimal strategy.
+So, for instance, a $\phi$ of 0.05 denotes the isorisk curve over $t$ where, at given $\delta(P, t)$, distancing is the preferable solution as long as its costs are less than, or equal to, 0.05 times that of illness or less.
 
@@
 
-As Figure~\ref{fig:fig3-SIR-by-delta} indicates for a given $R_0$ and $\frac{\beta}{\gamma}$, social distancing can have an overwhelmingly significant effect on the number of infectious individuals in a closed population, dependent on the number of persons in the population already engaged in social distancing.
+As Figure~\ref{fig:fig3-SIR-by-delta} indicates for empirically ascertained values of $R_0$ and $\gamma$ based on the literature on SARS-CoV-2,\cite{Li:2020aa,liu2020reproductive} social distancing can have an overwhelmingly significant effect on the number of infectious individuals in a closed population, and the magnitude of this effect is dependent on the number of persons in the population already engaged in social distancing.
 This effect is most pronounced early in the epidemic (approx.
 2-3 $\gamma^{-1}$ days), and the effect is most significant where less than half of the population is engaged in social distancing.
-Thus, unlike collective immunity in the case of vaccination, which often necessitates a fairly high level of penetration (typically estimated as $1 - R_0^{-1}$), social distancing can play a meaningful role in particular where much of the population is not engaged in such behaviours.
-This result can meaningfully inform a policy of encouraging individual social distancing early in an outbreak and dispelling perceptions that unless a critical volume of individuals are participating, marginal action is unnecessary.
+Thus, unlike collective immunity in the case of vaccination, which often necessitates a fairly high level of penetration (typically estimated as $1 - R_0^{-1}$), social distancing can play a meaningful role in particular where much of the population is not yet engaged in such behaviour.
+This result can meaningfully inform a policy of encouraging individual social distancing early in an outbreak and dispel the misperception that marginal action is unnecessary unless a critical volume of individuals are already participating.
 
@@
 
-Based on the key epidemiological dynamics data on the SARS-CoV-2 pandemic, the highest marginal effect of social distancing takes place in the same early timeframe of approx.
+Based on the key epidemiological dynamics data on the SARS-CoV-2 pandemic,\cite{Li:2020aa,liu2020reproductive} the highest marginal effect of social distancing takes place in the same early timeframe of approx.
 2-3 $\gamma^{-1}$ days.
-Unsurprisingly, even without integrating the time-dependent discount factor proposed by Reluga (2010), the numerical solutions indicate that the effect of social distancing is most significant where it is not yet a widely adopted strategy: for SARS-CoV-2, based on an initial population of 10,000 with a seed population of 0.1\% infected, the greatest marginal utility is encountered where less than 20\% of the population is engaged in such behaviour, and the effect is significantly less noticeable once $\delta(P)$ reaches 0.5.
+Unsurprisingly, even without integrating the time-dependent discount factor proposed by Reluga (2010),\cite{reluga2010game} the numerical solutions indicate that the effect of social distancing is most significant where it is not yet a widely adopted strategy: for SARS-CoV-2, based on an initial population of 10,000 with a seed population of 0.1\% infected, the greatest marginal utility is encountered where less than 20\% of the population is engaged in social distancing, and the effect is significantly less noticeable once $\delta(P, t)$ reaches 0.5 (Figure~\ref{fig:marginal_utility}).
 
-Calculations of marginal utility matter because they can guide public policy in determining what fraction of the population may feasibly be exempted from social distancing.
-Given the need for critical services to continue even in the throes of a pandemic, marginal utility calculations based on empirical data on an outbreak itself, as well as the societal response (as expressed by $\delta(P, t)$), may contribute to more accurate public health measures while limiting the effect of such measures on the economy.
+Calculations of marginal utility matter because they can guide public policy in determining what fraction of the population may feasibly be exempted from social distancing while still maintaining much of the benefit.
+Given the need for critical services, swuch as urgent medical care, emergency services and critical supply chain activities to continue even in the throes of a pandemic, marginal utility calculations based on empirical data on an outbreak may, along with the societal response (as expressed by $\delta(P, t)$, which can be empirically ascertained as well), contribute to more accurate public health measures while limiting the effect of such measures on the economy and on day-to-day life.
 
@@
 
-The costs both of social distancing and of failing to do so are notoriously difficult to quantify accurately.
+The accurate direct and indirect costs both of social distancing and of failing to do so are notoriously difficult to quantify accurately.
 Indeed, many of these costs are by their very nature not amenable to quantification.
-At the same time, by quantifying the relative ratio of cost of distancing ($c_d$) and cost of illness ($c_i$), we can identify a strategy-associated cost relationship that, given solid information, can assist in societal decision-making as to social distancing.
+At the same time, by quantifying the relative ratio of cost of distancing ($c_d$) and cost of illness ($c_i$), we can identify a strategy-associated cost ratio $\phi$ that, given approximations or empirical estimates of those costs, can assist in societal decision-making with regard to social distancing.
 As Figure~\ref{fig:cost_fraction} shows, for most cases, the cost of social distancing would have to exceed the cost of illness by at least an order of magnitude to make it a preferable strategy.
@@
 
-Estimates of direct medical costs of COVID-19 are difficult, but at least one figure puts the median cost of a symptomatic infection at US\$3,045 in direct costs alone,\cite{bartsch2020potential} typically compounded by loss of earnings, long-term physical harm, reduction in life expectancy and quality of life and, in severe cases, the risk of mortality.
+Estimates of direct medical costs of COVID-19 are difficult, but at least one study puts the median cost of a symptomatic infection at US\$3,045 in direct costs alone,\cite{bartsch2020potential} typically compounded by loss of earnings, long-term physical harm, reduction in life expectancy and quality of life and, in severe cases, the risk of mortality.
 The costs of social distancing are much less amenable to quantification, as these costs are primarily governed by indirect factors and intangibles, such as the cost of deferred medical treatment, second-order effects of the psychological burden inherent in decreased social interaction and the cost of lost revenue.
-While quantification, thus, of both the cost of distancing and cost of illness remains an outstanding subject of research, the cost fraction calculations can assist in reasoning about the best social strategy in view of these factors once ascertained.
+While quantification, thus, of both the cost of distancing and cost of illness remains an outstanding subject of research, the cost fraction calculations can assist in reasoning about the best social strategy in view of these factors once ascertained or estimated.
 
@@
 Pandemics pose a significant challenge to public health and social decision-making, and the COVID-19 pandemic is by no means an exception.
-Non-pharmaceutical interventions, such as social distancing, play a significant role in the public health arsenal and may play a determinative role in stemming the tide of an infectious disease that is otherwise not amenable to treatment or prophylaxis.
-Thus, until a vaccine or a reliable therapeutic, ideally with prophylactic properties, is found, non-pharmaceutical interventions are the mainstay of public health activity in the face of COVID-19.
+Non-pharmaceutical interventions, such as social distancing, play a significant role in the arsenal of tools that public health authorities can bring to bear on an epidemic that is otherwise not amenable to treatment or prophylaxis.
+Thus, until a vaccine or a reliable therapeutic, ideally with prophylactic properties, is found, non-pharmaceutical interventions are poised to remain the mainstay of public health activity in the face of COVID-19.
+In view of this, an increased understanding of the way NPIs that rely on social distancing affect the statistical dynamics of SARS-CoV-2 in a population is essential for sound decisin-making.
 
 This paper discussed a subject that is not devoid of controversy, both in the scientific and in the public realm.
-By their nature, non-pharmaceutical interventions interfere with citizens' day-to-day lives and may have complex economic, social and psychological effects.
+By their nature, NPIs interfere with citizens' day-to-day lives and may have complex economic, social and psychological effects.
 It is therefore important that strategy options are adequately explored from a quantitative perspective.
-It is hoped that in reinforcing the case for social distancing through an analysis of the statistical dynamics that underlie it, this paper can add to the wealth of literature in support of social distancing as an effective and cost-efficient non-pharmaceutical intervention where other tools are unavailable or inappropriate.
+It is hoped that in reinforcing the case for social distancing through an analysis of the statistical dynamics that underlie it, this paper can add to growing body of knowledge in support of social distancing as an effective and cost-efficient NPI where other tools are unavailable or inappropriate.
 
